### 2017-11-28 paper/sections/1_introduction.tex
paper: final draft
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

ParSplice, the molecular dynamics simulation, is representative of an important
class of HPC applications with similar working set behaviors that extensively
use software-based caches. It uses a hierarchy of caches and a single
persistent key-value store to store both observed minima across a molecule's
equation of motion (EOM) and the hundreds or thousands of partial trajectories
calculated each second during a parallel job.  This workload is pervasive
across simulations that (1) rely on a mesh-based decomposition of a physical
region and (2) result in millions or billions of mesh cells, where each cell
contains materials, pressures, temperatures and other characteristics that are
required to accurately simulate phenomena of interest.  The fine-grained data
annotation capabilities provided by key-value storage is a natural match for
these types of scientific simulations.  Unfortunately, simulations of this size
saturate the capacity and bandwidth capabilities of a single node so we need
more effective data management techniques.

>>>>

The ParSplice molecular dynamics simulation is representative of an important
class of HPC applications with similar working set behaviors that extensively
use software-based caches. It uses a hierarchy of caches and a single
persistent key-value store to store both observed minima across a molecule's
equation of motion (EOM) and the hundreds or thousands of partial trajectories
calculated each second during a parallel job.  This workload is pervasive
across simulations that (1) rely on a mesh-based decomposition of a physical
region and (2) result in millions or billions of mesh cells, where each cell
contains materials, pressures, temperatures and other characteristics that are
required to accurately simulate phenomena of interest.  The fine-grained data
annotation capabilities provided by key-value storage is a natural match for
these types of scientific simulations.  Unfortunately, simulations of this size
saturate the capacity and bandwidth capabilities of a single node so we need
more effective data management techniques.

### 2017-11-28 paper/sections/2_parsplice-keyspace-analysis.tex
paper: final draft
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\subsubsection{\underline{Initial conditions influence key activity}}
Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
pairs from the worker's cache for two different initial conditions of
\(\Delta\), which is the rate that new atoms enter the simulation.  The line is
the read request rate (\(y1\) axis) and the dots along the bottom are the
number of unique keys accessed (\(y2\) axis).  The \(\Delta_1\) growth rate
adds atoms every 100K microseconds while the \(\Delta_2\) growth rate adds
atoms every 1 million microseconds. So \(\Delta_2\) has a smaller growth rate
resulting in hotter keys and a smaller keyspace.  Values smaller than
\(\Delta_2\)'s growth rate or a temperature of 400 degrees result in very
little database activity because state transitions take too long. Similarly,
values larger than \(\Delta_1\)'s growth rate or a temperature of 4000 degrees
result in an equally meaningless simulation as transitions are unrealistic.
This figure demonstrates that small changes to \(\Delta\) can have a strong
effect on the timing and frequency with which new EOM minima are discovered and
referenced. This finding suggests that we need a flexible policy language and
engine to explore these trade-offs.

@@

\subsubsection{\underline{Entropy increases over time}} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time.  The number of read and
write requests are highest at the beginning of the run when tasks generate
segments for the same state, which is computationally cheap (this motivates
Section~\S\ref{sec:arch-specific}).  The resulting key access imbalance for the
two growth rates in Figure~\ref{fig:motivation-regimes} are shown in
Figure~\ref{fig:methodology-keys}, where reads are plotted for each unique
state, or key, along the \(x\) axis. Keys are more popular than others (up to
\(5\times\)) because worker tasks start generating states with different
coordinates later in the run.  Figure~\ref{fig:methodology-keys} shows that the
number of reads changes with different initial conditions (\(\Delta\)), but
that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
still \(5\times\) more popular than others).

Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
rates has temporal locality, as the reads per second for \(\Delta_2\) look like
the reads per second for \(\Delta_1\) stretched out along the time axis.
Trends also exist for temperature and number of workers but are omitted here
for space. This structure means that we can learn the key access patterns and
adapt the storage system accordingly (this motivates
Section~\S\ref{sec:dom-specific}).

>>>>

\subsubsection{\underline{Initial conditions influence key activity}}
Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
pairs from the worker's cache for two different initial conditions of
\(\Delta\), which is the rate that new atoms enter the simulation.  The line is
the read request rate (\(y1\) axis) and the dots along the bottom are the
number of unique keys accessed (\(y2\) axis).  The access patterns for
different growth rates have temporal locality, as the reads per second for
\(\Delta_2\) look like the reads per second for \(\Delta_1\) stretched out
along the time axis.  The \(\Delta_1\) growth rate adds atoms every 100K
microseconds while the \(\Delta_2\) growth rate adds atoms every 1 million
microseconds. So \(\Delta_2\) has a smaller growth rate resulting in hotter
keys and a smaller keyspace.  Values smaller than \(\Delta_2\)'s growth rate or
a temperature of 400 degrees result in very little database activity because
state transitions take too long. Similarly, values larger than \(\Delta_1\)'s
growth rate or a temperature of 4000 degrees result in an equally meaningless
simulation as transitions are unrealistic.

This figure demonstrates that small changes to \(\Delta\) can have a strong
effect on the timing and frequency with which new EOM minima are discovered and
referenced.  Trends also exist for temperature and number of workers but are
omitted here for space.  This finding suggests that we need a flexible policy
language and engine to explore these trade-offs.

@@

\subsubsection{\underline{Entropy increases over time}} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time.  The number of read and
write requests are highest at the beginning of the run when tasks generate
segments for the same state, which is computationally cheap (this motivates
Section~\S\ref{sec:arch-specific}).  The resulting key access imbalance for the
two growth rates in Figure~\ref{fig:motivation-regimes} are shown in
Figure~\ref{fig:methodology-keys}, where reads are plotted for each unique
state, or key, along the \(x\) axis. Keys are more popular than others (up to
\(5\times\)) because worker tasks start generating states with different
coordinates later in the run.  Figure~\ref{fig:methodology-keys} also shows that the
number of reads changes with different initial conditions (\(\Delta\)), but
that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
still \(5\times\) more popular than others).
